---
title: 'Lecture 11 - Normal and Normal Approximation'
author: 'Wesley Burr'
date: "05/12/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Snippet of Code from last Friday

```{r}
x <- seq(from = -3.5, to = 3.5, by = 0.01)
plot(x = x, y = dnorm(x), type = "l", 
     xlab = "z", ylab = "Normal Density")
pnorm(q = 20, mean = 15, sd = 4, lower.tail = FALSE)
```

# Examples for current Friday

## Example 1

Suppose we toss a fair coin 20 times. What is the probability of getting between 9 and 11 heads?

## Example 2

In a particular program at Trent, 60% of students are men and 40% are women. In a random sample of 50 students what is the probability that more than half are women?

## Example 3

"Skip the Dishes" finds that 70% of people order through them and ask for Chinese food, and 30% for Italian food. Last week 426 orders were made through the local, Peterborough, unit. Find the probability that at least 200 orders were for Italian food.


