---
title: 'Lecture 15'
output:
   ioslides_presentation:
     font-family: Lato Semibold
     font-import: http://fonts.googleapis.com/css?family=Lato
     widescreen: yes
     css: ../style.css
     fig_caption: yes
---
<style>
citation {
  <!-- font-size: 4px; -->
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- This is not explicitly in the OpenIntro slides - inverse problems -->

# Inverse Normal Problems

## The Inverse Problem

As we've seen in the previous, we often need to take $Z$-scores and find probabilities from them. Sometimes this is in a normal problem, sometimes an approximation, and so on.

What if you **wanted to go backward**? What would this look like?

## Inverse Problem Statement

What if I told you "the probability of this event happening is 0.5".
What would the $Z$-score of such a setup be?

$$
P(Z \leq z_0) = 0.5
$$

What's the unknown here? **$z_0$**!

## How do we solve for $z_0$?

* trial and error?
* use tables?
* **use R!**

## Using R to Find $z_0$

```{r}
qnorm(p = 0.5, mean = 0, sd = 1, lower.tail = TRUE)
```

Does this make sense? This is saying that $z_0 = 0$ has probability to its left of 0.5.

```{r}
pnorm(q = 0, mean = 0, sd = 1, lower.tail = TRUE)
```

## Checking Again

```{r, fig.height = 4}
x <- seq(from = -3, to = 3, by = 0.01)
plot(x, dnorm(x), type = "l", xlab = "Z")
abline(h = 0)
abline(v = 0, col = "red")
```

## Practice

What value of $z_0$ has $P(-z_0 \leq Z \leq z_0) = 0.3$?

## Practice

What value of $z_0$ has $P(-z_0 \leq Z \leq z_0) = 0.3$?

```{r, fig.height = 3, echo = FALSE}
x <- seq(from = -3, to = 3, by = 0.01)
plot(x, dnorm(x), type = "l", xlab = "Z", xaxt = 'n', 
     ylab = "Probability")
abline(h = 0)
abline(v = c(-1, 1), col = "red")
axis(side = 1, line = 0, at = c(-1, 1), label = c("-z0", "z0"))
```

So what do we actually need to run?

## Practice

```{r}
qnorm(p = 0.15, mean = 0, sd = 1, lower.tail = TRUE)
```

This is **-z0**, because we used 0.15 area to the left, and
**lower.tail = TRUE**.

```{r}
qnorm(p = 0.15, mean = 0, sd = 1, lower.tail = FALSE)
```

And this is **$z_0$**, because **lower.tail = FALSE**.

## More Practice

We're now going to solve a series of problems in all the variations of this that we can do.

## Problem Type 1: Simple Left

What value of $z_0$ has $P(-2 \leq Z \leq z_0) = 0.7$?

## Problem Type 2: Simple Right

What value of $z_0$ has $P(-z_0 \leq Z \leq 1.0) = 0.4$ ?

## Problem Type 3: Symmetric, known

What value of $z_0$ has $P(-z_0 \leq Z \leq z_0) = 0.4$ ?

## Problem Type 4: Symmetric, Indirect information

If the region from $-z_0$ to $z_0$ is 90\% of the area under
a standard normal curve, find $z_0$. 

## Problem Type 5: Not Symmetric, known

What value of $z_0$ has $P(-2\cdot z_0 \leq Z \leq z_0) = 0.6$ ?

## Problem Type 6: Not Symmetric, Indirect information

If the 90th percentile of your distribution is equal to
11, and the 5th percentile is equal to 4, find $\mu$
and $\sigma$.

## $\;$